A community-maintained standard library of population genetic simulations

### 2019-12-19 popSimManu.tex

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models},
the distribution of fitness effects \citep{Boyko:2008cr,kim2017inference,tataru2017inference,Fortier703918,Huang2019,Vecchyo770966},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247,Barroso2019},
and the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming even more important.
Assessing accuracy in population genetics is challenging in large part because
 ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods that are validated by a set of bespoke simulations produced by individual
groups. Beyond inefficiency from duplicated effort, such a situation leads
to decreased reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software
and select the best tool for a given task, thereby substantially increasing the reliability of population genetic inferences.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here
%RNG: Removed the sap. :-P Feel free to put it back if you like it better.

@@

For non-human genomes we have have implemented two demographic histories for
\emph{Drosophila melanogaster} and one each from \emph{A. thaliana} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample, as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For \emph{Arabidopsis thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes. Additionally, we have also implemented two generic models that comprise a piecewise constant size model and an isolation with migration (IM) model.

@@

\paragraph*{Multi-population demographic models}
As \texttt{stdpopsim} implements multi-population demographic models, we also
explored parameter estimation of population divergence parameters. In particular,
we simulated data under multi-population models for humans and \emph{Drosophila}
and then inferred  parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal by fitting an IM model
with constant population sizes and bi-directional migration \citep{hey2004im}. Our motivation for using
an IM model was to mimic the approach often used on empirical datasets, where a relatively
simple model is fit that may not reflect the true underlying demography.
For human models with more than two-populations (e.g. \cite{gutenkunst2009inferring})
this means that we are inferring parameters for a model which itself does
not match the model from which the data were generated (Figure
\ref{fig:IM_popn_human}A and B).

@@

\section*{Acknowledgements}
We thank the Probabilistic Modeling in Genomics conference organizers with making this collaboration possible.
Early on in the project we were helped by many people including XXXXXX.
CCK and KEL were funded under NIH Award R35GM119856.
JRA and ADK were funded under NIH Award R01GM117241.
FR and GG were supported by a Villum Young Investigator award (project no. 00025300).
DODV is funded by a UC MEXUS-CONACYT Collaborative Grant and a DGAPA-PAPIIT grant (PAPIIT-IA200620).

>>>>

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models},
the distribution of fitness effects \citep{kim2017inference,tataru2017inference},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247},
and the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming even more important.
Assessing accuracy in population genetics is challenging in large part because
 ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods that are validated by a set of bespoke simulations produced by individual
groups. Beyond inefficiency from duplicated effort, such a situation leads
to decreased reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software
and select the best tool for a given task, thereby substantially increasing the reliability of population genetic inferences.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here
%RNG: Removed the sap. :-P Feel free to put it back if you like it better.

@@

For non-human genomes we have have implemented two demographic histories for
\emph{Drosophila melanogaster} and one each from \emph{A. thaliana} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample, as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For \emph{Arabidopsis thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes.

@@

\paragraph*{Multi-population demographic models}
As \texttt{stdpopsim} implements multi-population demographic models, we also
explored parameter estimation of population divergence parameters. In particular,
we simulated data under multi-population models for humans and \emph{Drosophila}
and then inferred  parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal by fitting an isolation with migration (IM) model
with constant population sizes and bi-directional migration \citep{hey2004im}. Our motivation for using
an IM model was to mimic the approach often used on empirical datasets, where a relatively
simple model is fit that may not reflect the true underlying demography.
For human models with more than two-populations (e.g. \cite{gutenkunst2009inferring})
this means that we are inferring parameters for a model which itself does
not match the model from which the data were generated (Figure
\ref{fig:IM_popn_human}A and B).

@@

\section*{Acknowledgements}
We thank the Probabilistic Modeling in Genomics conference organizers with making this collaboration possible.
Early on in the project we were helped by many people including XXXXXX. 
CCK and KEL were funded under NIH Award R35GM119856. 
JRA and ADK were funded under NIH Award R01GM117241. 
FR and GG were supported by a Villum Young Investigator award (project no. 00025300).
